---
title: "Practical Learning Machine Course Project"
author: "Teo Chun Yong"
date: "July 23, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Background

Using devices such as Jawbone Up, Nike FuelBand, and Fitbit it is now possible to collect a large amount of data about personal activity relatively inexpensively. These type of devices are part of the quantified self movement - a group of enthusiasts who take measurements about themselves regularly to improve their health, to find patterns in their behavior, or because they are tech geeks. One thing that people regularly do is quantify how much of a particular activity they do, but they rarely quantify how well they do it. In this project, our goal will be to use data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways. More information is available from the website here: http://groupware.les.inf.puc-rio.br/har (see the section on the Weight Lifting Exercise Dataset).

# Data

The training data for this project are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv

The test data are available here:

https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv

The data for this project come from this source: http://groupware.les.inf.puc-rio.br/har. If you use the document you create for this class for any purpose please cite them as they have been very generous in allowing their data to be used for this kind of assignment.

#Project Goals

The goal of our project is to predict the manner in which they did the exercise. This is the "classe" variable in the training set. We also created a report describing how we built our model, how we used cross validation, and what we think the expected out of sample error is, and why we made the choices we did. We also use our prediction model to predict 20 different test cases. 

#Prepare the datasets

Load libraries and setup working directory
```{r echo=TRUE}
rm(list = ls(all = TRUE))

setwd('C:/Users/Joseph Teo Chun Yong/Documents/Practical_Learning_Machine/practicalmachinelearning/data')

library(caret)

trainingRaw <- read.csv(file="pml-training.csv", header=TRUE, as.is = TRUE, stringsAsFactors = FALSE, sep=',', na.strings=c('NA','','#DIV/0!'))
testingRaw <- read.csv(file="pml-testing.csv", header=TRUE, as.is = TRUE, stringsAsFactors = FALSE, sep=',', na.strings=c('NA','','#DIV/0!'))
trainingRaw$classe <- as.factor(trainingRaw$classe) 
```

#Examine the data
```{r echo=TRUE}
str(trainingRaw)
```

#Cleaning variables

After looking at all the variables of the sets, there are a lot of NA values or empty variables for the prediction. As we are going to compute the prediction only on the accelerometers values of belt, forearm, arm and dumbell. So, the non-accelerometer measures are discarded.
```{r echo=TRUE}
NAindex <- apply(trainingRaw,2,function(x) {sum(is.na(x))}) 
trainingRaw <- trainingRaw[,which(NAindex == 0)]
NAindex <- apply(testingRaw,2,function(x) {sum(is.na(x))}) 
testingRaw <- testingRaw[,which(NAindex == 0)]
```

#Preprocessing variables
```{r echo=TRUE}
v <- which(lapply(trainingRaw, class) %in% "numeric")

preObj <-preProcess(trainingRaw[,v],method=c('knnImpute', 'center', 'scale'))
trainLess1 <- predict(preObj, trainingRaw[,v])
trainLess1$classe <- trainingRaw$classe

testLess1 <-predict(preObj,testingRaw[,v])
```

#Removing the non zero variables

Removing the variables with values that are near zero as they have no meaning in the predictions
```{r echo=TRUE}
nzv <- nearZeroVar(trainLess1,saveMetrics=TRUE)
trainLess1 <- trainLess1[,nzv$nzv==FALSE]

nzv <- nearZeroVar(testLess1,saveMetrics=TRUE)
testLess1 <- testLess1[,nzv$nzv==FALSE]
```

#Create cross validation set

The training set is divided in two parts, one for training and the other for cross validation
```{r echo=TRUE}
set.seed(23051976)

inTrain = createDataPartition(trainLess1$classe, p = 3/4, list=FALSE)
training = trainLess1[inTrain,]
crossValidation = trainLess1[-inTrain,]
```

#Train model

Train model with random forest due to its highly accuracy rate. The model is build on a training set of 28 variables from the initial 160. Cross validation is used as train control method.

```{r echo=TRUE}
library(randomForest)
modFit <- train(classe ~., method="rf", data=training, trControl=trainControl(method='cv'), number=5, allowParallel=TRUE)
modFit
```

#Accuracy on training set and cross validation set

Following the computation on the accuracy of trainig and cross validation set

Training set:
```{r echo=TRUE}
trainingPred <- predict(modFit, training)
confusionMatrix(trainingPred, training$classe)
```

#Cross validation set
```{r echo=TRUE}
cvPred <- predict(modFit, crossValidation)
confusionMatrix(cvPred, crossValidation$classe)
```

#RESULTS

Predictions on the real testing set
```{r echo=TRUE}
testingPred <- predict(modFit, testLess1)
testingPred
```